---
title: "Blog Post 2: Women in the Workplace"
author: "Olivia Varricchione"
date: "2024-02-20"
categories: [analysis]
---

Introduction

The data set is from the Bureau of Labor Statistics and the Census Bureau about women in the workforce. The data was retrieved from the tidyTuesday github from the following link: https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-05. There are 2088 observations in the data set, 522 for each year from 2013 to 2016, and 12 variables. The variables that are of interest for this analysis are the year, wage_percent_of_male, major_category, and percent_female. This analysis will focus on what major categories have the highest female percentage and how the average female wage percentage of their male counterparts has changed. This analysis will also create a linear model that can be used to predict the wage_percent_of_male based on the year.

Primary Visualizations

```{r}
library(tidyverse)
jobs_gender <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")

percent_data <- jobs_gender |> select(year, major_category, percent_female) |>
  group_by(major_category) |>
  summarise(mean = mean(percent_female),
            sd = sd(percent_female),
            n_observations = n()) |>
  mutate(se = sd / sqrt(n_observations),
         lb_se = mean - se,
         ub_se = mean + se) |>
  mutate(major_category = fct_reorder(.f = major_category, .x = mean))

ggplot(data = percent_data, aes(x = major_category, y = mean)) +
  geom_errorbar(aes(ymin = lb_se, ymax = ub_se)) +
  geom_point() +
  theme_minimal() +
  coord_flip() +
  theme(axis.text.y = element_text(size = 6)) +
  labs(y = "Mean", x = "Major Category")
```

From this graph, we see that healthcare practitioners and sales/office have the highest percent of female workers at around 65% and 59% respectively. We also see that natural resources, construction, and maintenance has the lowest percent of females with about 5%.

```{r}
wage_data <- jobs_gender |> select(year, wage_percent_of_male) |>
  filter(!is.na(wage_percent_of_male)) |>
  group_by(year) |>
  summarise(mean = mean(wage_percent_of_male),
            sd = sd(wage_percent_of_male),
            n_observations = n()) |>
  mutate(se = sd / sqrt(n_observations),
         lb_se = mean - se,
         ub_se = mean + se)

ggplot(data = wage_data, aes(x = year, y = mean)) +
  geom_errorbar(aes(ymin = lb_se, ymax = ub_se)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Year", y = "Mean wage percent of male")
```

From this visualization, we see that there has been some increase in the mean wage percent of male from 2013 to 2016. However, we also see that there is a fair amount of variability within the data, likely due to the differences across occupations.

```{r}
library(broom)
linear_mod <- lm(wage_percent_of_male ~ year, data = jobs_gender)
linear_mod |> tidy()
aug_mod <- linear_mod |> augment()
aug_mod
```

Checking assumptions:

Linearity:

```{r}
ggplot(data = aug_mod, aes(x = year, y = wage_percent_of_male)) +
  geom_jitter(width = 0.2, alpha = 0.2)
```

Normality:

```{r}
ggplot(data = aug_mod, aes(x = .resid)) +
  geom_histogram(bins = 30, color = "black", fill = "white")
```

Constant Variance:

```{r}
ggplot(data = aug_mod, aes(x = .fitted, y = .resid)) +
  geom_jitter(width = 0.2, alpha = 0.2)
```

Visualizing the model:

```{r}
grid <- tibble::tibble(year = 2013:2016)

aug_linear <- augment(linear_mod, newdata = grid, interval = "confidence")

ggplot(aug_linear, aes(x = year, y = .fitted), width = 0.25) +
  geom_point() +
  geom_line(alpha = 0.4) +
  geom_errorbar(aes(ymin = .lower, ymax = .upper)) +
  theme_minimal() +
  labs(caption = "Error bars are 95% confidence intervals for the mean")
```

From the visualization, we see that there are increasing predictions of the mean wage percent of male through the years. From the corresponding table of coefficients, we see that for every year increase the model predicts that the wage percent of male goes up by 0.201.

Conclusion and Wrap-Up

The major category of healthcare practitioners had the highest percent of female workers while natural resources, construction, and maintenance had the lowest percentage of female workers. The wage percent of male increased from 2013 to 2016 and the model predicted that for every year the predicted wage percent of male goes up by 0.201. In the future, it would be interesting to include more predictors in the model, to see how other factors influence the predicted wage percent of male.

Connection to Class Ideas

The first two visualizations connect to class because of the inclusion of errorbars. In chapter 7 we discussed the importance of showing variability in the visualizations and how to include error bars in plots. These ideas were implemented when creating the plots to answer the first two questions. Additionally, the model connects to class since chapter 8 was reviewing concepts from STAT 213, focusing on how to interpret the coefficients and how to make these plots in R. The 4 step strategy from chapter 8 was implemented to visualize the model predictions.
