---
title: "Blog Post 3: Chocolate Rating"
author: "Olivia Varricchione"
date: "2024-03-2"
categories: [analysis]
---

Introduction

This analysis will be using the chocolate data set from tidyTuesday: https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md which has 2530 observations. The variables of interest include rating, cocoa percent, country of bean origin, and company location. More specifically, the country of bean origin and company location were used to create variables for the continent of bean origin and continent of company location to decrease the different levels of those variables. This analysis will focus on predicting the chocolate rating based on the cocoa percent, the continent of bean origin, and the company continent.

Primary Visualizations

To start, the data was wrangled to change the country of bean origin and the company country to be in the form of a continent so it would be a predictor that is much easier to visualize since it has less levels.
```{r}
chocolate <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')
library(tidyverse)
library(broom)
chocolate_data <- chocolate |> 
  select(rating, cocoa_percent, country_of_bean_origin, company_location) |> 
  separate(cocoa_percent, into = "cocoa_percent", sep = -1) |> 
  mutate(cocoa_percent = as.numeric(cocoa_percent)) |>
  mutate(company_continent = case_when(
    company_location %in% c("Canada", "Costa Rica", "Dominican Republic", "El Salvador", "Grenada", "Guatemala", "Honduras", "Martinique", "Mexico", "Nicaragua", "Puerto Rico", "St. Lucia", "St. Vincent-Grenadines", "U.S.A.") ~"North America",
    company_location %in% c("Argentina", "Bolivia", "Brazil", "Chile", "Colombia", "Ecuador", "Peru", "Suriname", "Venezuela") ~"South America",
    company_location %in% c("Amsterdam", "Austria", "Belgium", "Czech Republic", "Denmark", "Finland", "France", "Germany", "Hungary", "Iceland", "Ireland", "Italy", "Lithuania", "Netherlands", "Norway", "Poland", "Portugal", "Russia", "Scotland", "Spain", "Sweden", "Switzerland", "U.K.", "Wales") ~"Europe",
    company_location %in% c("Ghana", "Madagascar", "Sao Tome", "Sao Tome & Principe", "South Africa") ~"Africa",
    company_location %in% c("India", "Israel", "Japan", "Malaysia", "Philippines", "Singapore", "South Korea", "Taiwan", "Thailand", "U.A.E.", "Vietnam") ~"Asia",
    company_location %in% c("Australia", "Fiji", "New Zealand", "Vanuatu") ~"Oceania")) |>
  mutate(bean_continent = case_when(
    country_of_bean_origin %in% c("Belize","Costa Rica", "Cuba", "Dominican Republic", "El Salvador", "Grenada", "Guatemala", "Haiti", "Honduras", "Jamaica", "Martinique", "Mexico", "Nicaragua", "Panama", "Puerto Rico", "St. Lucia", "St. Vincent-Grenadines", "U.S.A.") ~ "North America",
    country_of_bean_origin %in% c("Bolivia", "Brazil", "Colombia", "Ecuador", "Peru", "Suriname", "Tobago", "Trinidad", "Venezuela") ~ "South America",
    country_of_bean_origin %in% c() ~ "Europe",
    country_of_bean_origin %in% c("Cameroon", "Congo", "DR Congo", "Gabon", "Ghana", "Ivory Coast", "Liberia", "Madagascar", "Nigeria", "Principe", "Sao Tome", "Sao Tome & Principe", "Sierra Leone", "Tanzania", "Togo", "Uganda") ~ "Africa",
    country_of_bean_origin %in% c("Burma", "China", "India", "Indonesia", "Malaysia", "Philippines", "Sri Lanka", "Sulawesi", "Sumatra", "Taiwan", "Thailand", "Vietnam") ~ "Asia",
    country_of_bean_origin %in% c("Australia", "Fiji", "Papua New Guinea", "Samoa", "Solomon Islands", "Vanuatu") ~ "Oceania")) |>
  mutate(bean_continent = as.factor(bean_continent)) |>
  mutate(company_continent = as.factor(company_continent)) |>
  na.omit()
```

A linear model was then created to predict the chocolate rating based on the cocoa percent, continent of bean origin, and the company continent.
```{r}
model <- lm(rating ~ cocoa_percent + bean_continent + company_continent, data = chocolate_data)
model |> tidy()
```
We can see that for every increase in the cocoa percentage, the predicted rating is expected to decrease by 0.0112 assuming the continent of bean origin and company continent remains the same. Similarly, we see that for a bean origin of Asia, North America, and Oceania the predicted rating is expected to be less than the reference (Africa) assuming the cocoa percentage and company continent remains the same. However, for a South America bean origin, the predicted rating is expected to be 0.00518 higher than Africa for the same cocoa_percent and company continent. Finally, we see that for a company continent of Asia, Europe, North America, Oceania, or Sourth America the predicted rating is expected to be higher than the reference (Africa) if the cocoa percentage and bean origin remain constant. For example, for a European company, the rating is predicted to be 0.211 higher than an African company for the same cocoa percentage and the same bean origin continent. All of these observations would be easier to see in a visualization. So, let's visualize this data.

```{r}
library(modelr)
grid <- chocolate_data |>
  data_grid(
    cocoa_percent = seq_range(cocoa_percent, n = 6),
    company_continent = chocolate_data |> pull(company_continent) |> levels(),
    bean_continent = chocolate_data |> pull(bean_continent) |> levels()
  )
grid

aug <- augment(model, newdata = grid, interval = "confidence")
aug

ggplot(data = chocolate_data, aes(x = cocoa_percent, y = rating)) +
  geom_point(alpha = 0.2) +
  geom_line(data = aug, aes(y = .fitted, colour = company_continent)) +
  facet_wrap(~bean_continent) +
  theme_minimal() +
  labs(y = "Predicted Rating", x = "Cocoa Percent") +
  scale_colour_viridis_d()
```

From this visualization, we can see that for all of the company continents, the predicted rating decreases for higher cocoa percents. We can also see that African companies seem to have a lower predicted rating whereas Oceania companies seem to have a higher rating. However, we can also see that this might not be the best model as the relationship, meaning the slope, between cocoa percent and rating does not change depending on the company continent. We can fix this by incorporating an interaction between variables.

```{r}
model2 <- lm(rating ~ cocoa_percent + bean_continent + cocoa_percent:bean_continent + company_continent, data = chocolate_data)
model2 |> tidy()

aug2 <- augment(model2, newdata = grid, interval = "confidence")
aug2

ggplot(data = chocolate_data, aes(x = cocoa_percent, y = rating)) +
  geom_point(alpha = 0.2) +
  geom_line(data = aug2, aes(y = .fitted, colour = company_continent)) +
  facet_wrap(~bean_continent) +
  theme_minimal() +
  scale_colour_viridis_d() +
  labs(y = "Predicted Rating", x = "Cocoa Percent")
```

From this visualization we can see that if the bean continent is Africa or South America, there is a steeper slope in the decrease in rating as the cocoa percent increases, whereas Oceania does not have too much of a decrease with an increase in cocoa percentage.

Conclusion and Wrap-Up

From this analysis found that the as the cocoa percentage increased, the predicted rating decreased, with African companies having lower ratings and Oceania companies having higher ratings. Additionally, there was a steeper decrease for African and South American bean origins and a more level trend for Oceania bean origins. In the future, it would be interesting to see the trend of individual countries within the continents, as this is the data that was originally provided and it would allow us to see more specifically which countries have the best retention of rating with an increase in cocoa percentage.


Connection to Class Ideas

These visualizations relate to class because of the use of linear models. We have spent a lot of time in class using the model, grid, augment, visualize method to create data visualizations for models. These visualizations makes it easier to see the trends in predicted values, especially if the target audience is someone without a statistical background.